### 2019-06-19 paper-three/paper-three.tex
initial notes on paper three

TODO: Attention in one place only

>>>>

Mapping all of sensory experience to these eight categories begs the question of
attention, of awareness: Where does the meditator try to fasten her awareness? Where
is awareness normally? For the average person, awareness jumps around across these
eight categories. Even when one tries to focus on a difficult intellectual problem,
the discomforts of back pain and hunger or the distraction of an irritating sound
would draw attention away from thought, the desired object of attention. Vipassana
meditation asks the meditator to use bodily sensation as a gateway to the other seven
sense experiences. Rather than focusing on sound, focus on the sensation generated in
the body by the ear sense door. Rather than focusing on a thought or emotion, focus
on the sensation in the body generated by that thought or emotion. This is extremely
difficult to do, which is why (for lay people, in most cases) a 10-day silent
residential course \cite{dhamma} is necessary to learn the technique.

\begin{center}
  \Huge{The Mundane and The Supramundane}
\end{center}

Reasons for meditating will be broken down into two categories. \textit{The Mudane}
in this context refers not to the tedious but to the earthly, the material. Most people will begin
meditating for reasons in the mundane field simply because most people have never
experienced the supramundane. \textit{The Supramundane} in this context refers not to
spirituality or religion but experiences which transcend the material, physical
world.

Because supramundane experiences tend to occur only in deep meditative states,
the reasons for meditating listed will predominantly fall in the mundane
category. Whether or not the meditator experiences supramundane states in deep
meditation, these altered states of mind are never the goal of meditation. The goal
of meditation is to change the meditator's mental habits, to move away from unhealthy
mental patterns which cause harmful behaviours toward healthy mental patterns which
encourage productive behaviour. Obviously this change is only visible in the mundane
world, outside of meditation.

\begin{center}
  \Huge{Meditation vs. Naps}
\end{center}

While staying at a friend's house, I excused myself in the evening to meditate. He
sincerely asked, ``Is meditating for an hour really more valuable than using that
time for a good nap?''
